---
title: "Networks"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    includes:
      before_body: [includes/include_header.html, includes/include_header_navpage.html]
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

# Networks

Network diagrams, or graph, are extremely useful visualisations to communicate the connections between distinct elements in your data; for instance the collaborations between researchers in a particular field or the connections between transport hubs in a logistics network. All network diagrams require the same basic data: nodes, edges and a layout algorithm.

### Nodes

Nodes are the unique elements in your data (the individuals). They might well have a number of different properties which are important when visualising the data:

- Node appearance: nodes are often styled to convey information about the thing the node represents, typical channels of communication include size, colour and shape
- Intrinsic properties: nodes may have intrinsic properties that are important if one was constructing a journey (or path) through the network and might be displayed to the user when nodes are clicked or hovered over.

It's extremely important to understand that **in general the absolute positions of nodes is unimportant**. Most [layout algorithms](networks.html#layout-algorithms) are non-deterministic, i.e. final node positions are computed by running a computer simulation. The three graph below display exactly the same data using a very general purpose layout algorithm but with slightly different random seeds (initial positions of the nodes for the simulation):

<!--html_preserve-->
<div class="row">
   <div class="col-md-4">

```{r}
library("tidyverse")
library("visNetwork")

my_nodes <- tribble(
  ~id, ~label, ~color, ~size,
  1, "Character A", "purple", 10, 
  2, "Character B", "purple", 13,
  3, "Character C", "orange", 17, 
  4, "Character D", "orange", 12,
  5, "Character E", "purple", 17,
  6, "Character Z", "green", 10
)

my_edges <- tribble(
  ~from, ~to,
  1, 2,
  3, 1,
  4, 1,
  4, 2,
  5, 2,
  5, 3,
  5, 6
)
visNetwork(my_nodes,
           my_edges,
           width = "200px",
           height = "200px") %>%
  visIgraphLayout(randomSeed = 1) %>%
  visInteraction(zoomView = FALSE)
```

</div>

   <div class="col-md-4">

```{r}
visNetwork(my_nodes,
           my_edges,
           width = "200px",
           height = "200px") %>%
  visIgraphLayout(randomSeed = 15) %>%
  visInteraction(zoomView = FALSE)
```

</div>

   <div class="col-md-4">

```{r}
visNetwork(my_nodes,
           my_edges,
           width = "200px",
           height = "200px") %>%
  visIgraphLayout(randomSeed = 2002) %>%
  visInteraction(zoomView = FALSE)
```

</div>
</div>
<!--/html_preserve-->

### Edges

Edges communicate how two nodes are connected to one another, they have the following properties:

- Directed/Undirected: An undirected edge indicates there's no direction in the relationship between two nodes, whereas a directed node indicates information can only pass through the relation between these two nodes in the direction specified. Note that directed nodes might well be bi-directional.
- Edge appearance: edges are often styled to convey information about the relationship between the nodes, typical channels of communication include width, dashed/dotted and colour.
- Intrinsic properties: edges may have intrinsic properties that are important if one was constructing a journey (or path) through the network.

The networks below are undirected and directed, respectively. Select a node to see it's first degree neighbours, note that in the second graph the direction of the edges affects which nodes are highlighted.

<!--html_preserve-->
<div class="row">
   <div class="col-md-6">
<center>
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library("tidyverse")
library("visNetwork")
library("igraph")

my_nodes <- tribble(
  ~id, ~label, ~color, ~size,
  1, "Character A", "purple", 10,
  2, "Character B", "purple", 13,
  3, "Character C", "orange", 17,
  4, "Character D", "orange", 12,
  5, "Character E", "purple", 17,
  6, "Character Z", "green", 10
)

my_edges <- tribble(
  ~from, ~to, ~width,
  1, 2, 5,
  # 2, 1, 5,
  3, 1, 1,
  # 1, 3, 1,
  4, 1, 6,
  # 1, 4, 6,
  4, 2, 3,
  # 2, 4, 3,
  5, 2, 2,
  # 2, 5, 2,
  5, 3, 5,
  # 3, 5, 5,
  5, 6, 5
  # 6, 5, 5
)

visNetwork(my_nodes,
           my_edges,
           width = "200px",
           height = "200px") %>%
  visOptions(highlightNearest = list(enabled = TRUE, degree = 1)) %>%
  visInteraction(zoomView = FALSE)
```
</center>
</div>

<div class="col-md-6">
<center>
```{r}
library("tidyverse")
library("visNetwork")

my_nodes <- tribble(
  ~id, ~label, ~color, ~size,
  1, "Character A", "purple", 10, 
  2, "Character B", "purple", 13,
  3, "Character C", "orange", 17, 
  4, "Character D", "orange", 12,
  5, "Character E", "purple", 17,
  6, "Character Z", "green", 10
)

my_edges <- tribble(
  ~from, ~to, ~width,
  1, 2, 5,
  3, 1, 1,
  4, 1, 6,
  4, 2, 3,
  5, 2, 2,
  5, 3, 5,
  5, 6, 5
)
visNetwork(my_nodes,
           my_edges,
           width = "200px",
           height = "200px") %>%
  visIgraphLayout(randomSeed = 1) %>%
  visEdges(arrows = list(to = list(enabled = TRUE, scaleFactor = 1))) %>%
  visOptions(highlightNearest = list(enabled = TRUE, algorithm =  "hierarchical", degree = list(from = 0, to = 1))) %>%
  visInteraction(zoomView = FALSE)
```
</center>
</div>
</div>
<!--/html_preserve-->

### Layout algorithms {#layout-algorithms}

As mentioned above, **in general the absolute positions of nodes is unimportant**. Most [layout algorithms](networks.html#layout-algorithms) are non-deterministic, i.e. final node positions are computed by running a computer simulation. Deciding on which layout algorithm (or graph embedding) to use for your visualisations is somewhat of an art as opposed to a science. The three graph below display the exact same dataset using three very different layout algorithms from the excellent [igraph library](http://igraph.org/r/doc/layout_.html){target='_blank'}. In the future this website will include a page dedicated to choosing layout algorithms.

<!--html_preserve-->
<div class="row">
   <div class="col-md-4">
<center>

```{r}
library("igraph")
my_random_igraph <- erdos.renyi.game(100, 1/39)
my_random_igraph_components <- decompose(my_random_igraph)
component_vcounts <- sapply(my_random_igraph_components, function(x) vcount(x))
largest_component <- my_random_igraph_components[[which(component_vcounts == max(component_vcounts))]]

random_visN <- toVisNetworkData(largest_component)

visNetwork(random_visN$nodes,
           random_visN$edges,
           width = "200px",
           height = "200px",
           main = "layout_in_circle") %>%
  visIgraphLayout(layout = "layout_in_circle")
```
</center>
</div>
   <div class="col-md-4">
<center>
```{r}
visNetwork(random_visN$nodes,
           random_visN$edges,
           width = "200px",
           height = "200px",
           main = "layout_nicely") %>%
  visIgraphLayout(layout = "layout_nicely")
```

</center>
</div>
   <div class="col-md-4">
<center>
```{r}
visNetwork(random_visN$nodes,
           random_visN$edges,
           width = "200px",
           height = "200px",
           main = "layout.grid") %>%
  visIgraphLayout(layout = "layout_nicely")
```

</center>
</div>
</div>
<!--/html_preserve-->

