---
title: "Other: Diagrams and Flowcharts"
output:
  html_document:
    toc: false
    toc_float:
      collapsed: false
    includes:
      before_body: [includes/include_header.html, includes/include_header_navpage.html]
editor_options: 
  chunk_output_type: console
---
# Diagrams and Flowcharts

Diagrams and flowcharts cover a wide range of so-called "technical drawings" and project management visualisations. These visualisations are often useful for explaining how a process works or to display the relationships between assets in a project. Interactivity may include the ability to select a specific path in the chart or to exclude a node. It's important to remember that designing useful diagrams/flowcharts is often a very time intensive process.

Required Data:

- Nodes of the diagram: Similarly to graph networks, diagrams consist of nodes that represent unique objects (or potentially outcomes) in the diagram
- Edges of the diagram: Similarly to graph networks, diagrams consist of edges that represent the connection (or flow) between nodes.

Here's an example flowchart that **Oxshef: dataviz** uses to advise which visualisation tool to choose:

```{r tool-selection-css, results="asis", echo=FALSE}
cat("
<style>
#mermaidChart0 { 
    min-height: 900px;
    min-width: 800px;
}
</style>
")
```

```{r tool-selection-flowchart, echo = FALSE, fig.height=6}
library("DiagrammeR")

oxshef_flowchart <- DiagrammeR("graph TB;
           A{Is your data public?} -- yes -->doiQuestion;
           A -- no -->B[<center><b>Oxshef: dataviz</b> exclusively supports researchers in <br> the visualisation and promotion of <br> public datasets.</center>];
           doiQuestion{<center>Is your dataset published<br> to a DOI-issuing data repository?</center>};
           doiNotIssued[<center>Please ensure to deposit your data<br> in a DOI-issuing repository</center>]
           doiQuestion -- yes --> externalDataQuestion;
           doiQuestion -- no --> doiNotIssued;
           externalDataQuestion{<center>Does your visualisation tool <br> allow data to be pulled from <br> your data repository?</center>};
           externalDataSupported[<center>Excellent!<br> Our resources are designed for you!</center>];
           externalDataUnsupported[<center>We recommend <b>not</b> using these tools</center>];
           externalDataUnknown[<center>Refer to our <br>http://oxshef-charts.netlify.com/tools.html page</center>];
           externalDataQuestion -- yes -->externalDataSupported;
           externalDataQuestion -- no -->externalDataUnsupported;
           externalDataQuestion -- I don't know -->externalDataUnknown",
           width = "100%",
           height = "900px")
library("htmltools")

div(oxshef_flowchart)
```

